### 2014-07-02 disclosure_game.tex

\documentclass{article}
\usepackage{ifthen}
\newboolean{enable-backrefs} % enable backrefs in the bibliography
\setboolean{enable-backrefs}{true} % true false
\newcommand{\backrefnotcitedstring}{\relax}%(Not cited.)
\newcommand{\backrefcitedsinglestring}[1]{(Cited on page~#1.)}
\newcommand{\backrefcitedmultistring}[1]{(Cited on pages~#1.)}
\ifthenelse{\boolean{enable-backrefs}}%
{%
		\PassOptionsToPackage{hyperpageref}{backref}
		\usepackage{backref} % to be loaded after hyperref package
		   \renewcommand{\backreftwosep}{ and~} % separate 2 pages
		   \renewcommand{\backreflastsep}{, and~} % separate last of longer list
		   \renewcommand*{\backref}[1]{}  % disable standard
		   \renewcommand*{\backrefalt}[4]{% detailed backref
		      \ifcase #1 %
		         \backrefnotcitedstring%
		      \or%
		         \backrefcitedsinglestring{#2}%
		      \else%
		         \backrefcitedmultistring{#2}%
		      \fi}%
}{\relax}
\usepackage{fullpage}
\usepackage{comment}
\usepackage[square, numbers]{natbib}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[subrefformat=parens,labelformat=parens]{subfig}
\usepackage{acronym}
\usepackage{nomencl}
\usepackage[style=long,nolist]{glossaries}
% the following is useful when we have the old nomencl.sty package
\providecommand{\printnomenclature}{\printglossary}
\providecommand{\makenomenclature}{\makeglossary}
\title{Deciding to Disclose: Pregnancy and Alcohol Misuse}
\author{Jonathan Gray \and Jakub Bijak \and Seth Bullock\\
University of Southampton, Southampton, SO17 1BJ, United Kingdom\\
E-mail: \texttt{j.gray@soton.ac.uk}}
\date{\vspace{-5ex}}
\begin{document}
\maketitle

%*************************
% Acronyms
%*************************
%\phantomsection
%\refstepcounter{dummy}
\pdfbookmark[1]{Acronyms}{acronyms}
%\markboth{\spacedlowsmallcaps{Acronyms}}{\spacedlowsmallcaps{Acronyms}}
\section*{Acronyms}
\begin{acronym}[UML]
  \acro{FAS}{Foetal Alcohol Syndrome}
  \acro{CPT}{Cumulative Prospect Theory}
  \acro{PT}{Prospect Theory}
  \acro{NICE}{National Institute for Health and Care Excellence}
  \acro{IAPT}{Improving Access to Psychological Therapies}
  \acro{CMACE}{Centre for Maternal and Child Enquiries}
  \acroindefinite{CMACE}{the}{the}
  \acro{AUDIT}{Alcohol Use Disorders Identification Test}
  \acro{T-ACE}{Tolerance, Annoyance, Cut down, Eye-opener}
  \acro{RCOG}{Royal College of Obstetricians and Gynecologists}
  \acroindefinite{RCOG}{the}{the}
  \acro{AML}{acute myeloid leukemia}
  \acro{ADHD}{Attention Defecit Hyperactivity Disorder}
  \acro{PND}{Postnatal Depression}
  \acro{RCT}{randomised control trial}
  \acro{WHO}{the World Health Organisation}
  \acro{ESS}{Evolutionarily Stable Strategy}
  \acro{MCMC}{Markov chain monte carlo methods}
  \acro{DU}{Discounted Utility}
  \acro{AIC}{Akaike information criterion}
  \acro{ANOVA}{analysis of variance}
  \acro{ALSPAC}{Avon Longitudinal Study of Parents and Children}
  \acro{GEM}{Gaussian Emulation Machine}
  \acrodefplural{GEM}{Gaussian Emulation Machines}
  \acro{GEM-SA}{Gaussian Emulation Machine for Sensitivity Analysis}
  \acro{JIT}{Just-in-time}
  \acro{ABM}{Agent Based Model}
  \acro{CPT}{Cumulative Prospect Theory}
\end{acronym}

\newglossaryentry{gravid}{name={gravid}, description={Pregnant}}
\newglossaryentry{term}{name={term}, description={Gestation of 37-42 weeks, 40 weeks is considered full term}}
\newglossaryentry{binge}{name={binge drinking}, description={Drinking with the express purpose of becoming drunk}}
\newglossaryentry{burnout}{name={burnout}, description={Lasting physical and emotional exhaustion, and disillusionment}}
\newglossaryentry{antenatal}{name={antenatal},description={The time period covering the pregnancy, prior to birth}}
\newglossaryentry{episiotomies}{name={episiotomies}, description={An episiotomy is a common procedure where an incision is made in the perineum to facilitate delivery, and attempt to reduce the risk of vaginal tearing}}
\newglossaryentry{neonatal asphyxia}{name={antenatal},description={The time period covering the pregnancy, prior to birth}}

%\refstepcounter{dummy}
\pdfbookmark[1]{Glossary}{glossary}
\printglossary[type=main]

\printnomenclature{}

\include{Abstract}

\section{Introduction} %(half page)

\include{Background}

\section{Model}

\subsection{Disclosure Game}

\subsection{Agent Models}

\section{Method}
\subsection{Qualitative Trends}
\subsection{Sensitivity Analysis}

\section{Results}
\subsection{}

\section{Discussion and Conclusions}

\bibliographystyle{plainnat}
\bibliography{library}
\end{document}