### 2014-08-26 SensitivityAnalysis.tex
Figures!

%!TEX root = disclosure_game.tex
%Include the descent into chaos bit
\section{Sensitivity Analysis}
\label{app:sensitivity_results}

\subsection{Lexicographic}

\subsubsection{Median Moderate Drinker Signalling}

\subsubsection{Median Between Groups IQR}

\subsubsection{Median Moderate Drinker Signalling IQR}

\subsubsection{Between Groups IQR IQR}