### 2014-08-30 SensitivityAnalysis.tex
Appendix

%!TEX root = disclosure_game.tex
%Include the descent into chaos bit
\section{Sensitivity Analysis}
\label{app:sensitivity_results}

\subsubsection{Median Moderate Drinker Signalling}

\begin{table}[h!]
\center
\begin{tabular} {|l | l | l | l | l | l |}
\hline
Parameter & Description & Lexicographic & Bayesian Payoff & Bayesian & \ac(CPT) \\ \hline
\(p_{w}(m)\) & Proportion of moderate drinkers & 0.367 & 1.145 \\ \hline
\(p_{w}(l)\) & Proportion of light drinkers & 10.080 & 37.750 \\ \hline
\(p_{m}(m)\) & Proportion of moderate midwives & 6.715 & 13.017\\ \hline
\(p_{m}(l)\) & Proportion of non-judgemental midwives & 43.942 & 1.655 \\ \hline
\(q_{w}\) & Probability of women sharing & 0.198 & 5.527 \\ \hline
\(w_{w}\) & Weight of shared information for women & 0.355 & 13.025 \\ \hline
\(q_{m}\) & Probability of midwives sharing & 0.145 & 0.667 \\ \hline
\(w_{m}\) & Weight of shared information for midwives & 0.118 & 0.376 \\ \hline
\(x_{h}\) & Health payoff for healthy delivery & 0.457 & 9.618 \\ \hline
\(s_{i}[a_{i}]:s_{i}[a_{\neg i}]\) & Psuedo-count favouring honesty & 0.140 & 7.537 \\ \hline
\end{tabular}
\caption[Table caption text]{Median moderate drinker signalling parameter sensitivity. \label{tab:sa_results_sig}}
\end{table}

\begin{table}[h!]
\center
\begin{tabular} {|l | l | l | l | l | l |}
\hline
Rule & \(\sigma^2\) & Nugget \(\sigma^2\) & Mean output & Total output variance & Code uncertainty & RMSSE \\ \hline
Lexicographic & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
Bayesian Payoff & 1.667 & 0.475 &  0.662 & 0.003 & 0.181 & 3.12 \\ \hline
Bayesian & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
\ac{CPT} & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
\end{tabular}
\caption[Table caption text]{Median moderate drinker signalling emulator statistics. \label{tab:sa_emulator_sig}}
\end{table}

\subsubsection{Median Between Groups IQR}

\begin{table}[h!]
\center
\begin{tabular} {|l | l | l | l | l | l |}
\hline
Parameter & Description & Lexicographic & Bayesian Payoff & Bayesian & \ac(CPT) \\ \hline
\(p_{w}(m)\) & Proportion of moderate drinkers & 0.367 & 1 \\ \hline
\(p_{w}(l)\) & Proportion of light drinkers & 10.080 & 1 \\ \hline
\(p_{m}(m)\) & Proportion of moderate midwives & 6.715 & 1 \\ \hline
\(p_{m}(l)\) & Proportion of non-judgemental midwives & 43.942 & 1 \\ \hline
\(q_{w}\) & Probability of women sharing & 0.198 & 1 \\ \hline
\(w_{w}\) & Weight of shared information for women & 0.355 & 1 \\ \hline
\(q_{m}\) & Probability of midwives sharing & 0.145 & 1 \\ \hline
\(w_{m}\) & Weight of shared information for midwives & 0.118 & 1 \\ \hline
\(x_{h}\) & Health payoff for healthy delivery & 0.457 & 100 \\ \hline
\(s_{i}[a_{i}]:s_{i}[a_{\neg i}]\) & Psuedo-count favouring honesty & 0.140 & 100:1 \\ \hline
\end{tabular}
\caption[Table caption text]{Median between groups IQR parameter sensitivity. \label{tab:sa_results_iqr}}
\end{table}

\begin{table}[h!]
\center
\begin{tabular} {|l | l | l | l | l | l |}
\hline
Rule & \(\sigma^2\) & Nugget \(\sigma^2\) & Mean output & Total output variance & Code uncertainty & RMSSE \\ \hline
Lexicographic & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
Bayesian Payoff & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
Bayesian & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
\ac{CPT} & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
\end{tabular}
\caption[Table caption text]{Median between groups IQR emulator statistics. \label{tab:sa_emulator_iqr}}
\end{table}

\subsubsection{Median Moderate Drinker Signalling IQR}

\begin{table}[h!]
\center
\begin{tabular} {|l | l | l | l | l | l |}
\hline
Parameter & Description & Lexicographic & Bayesian Payoff & Bayesian & \ac(CPT) \\ \hline
\(p_{w}(m)\) & Proportion of moderate drinkers & 0.367 & 1 \\ \hline
\(p_{w}(l)\) & Proportion of light drinkers & 10.080 & 1 \\ \hline
\(p_{m}(m)\) & Proportion of moderate midwives & 6.715 & 1 \\ \hline
\(p_{m}(l)\) & Proportion of non-judgemental midwives & 43.942 & 1 \\ \hline
\(q_{w}\) & Probability of women sharing & 0.198 & 1 \\ \hline
\(w_{w}\) & Weight of shared information for women & 0.355 & 1 \\ \hline
\(q_{m}\) & Probability of midwives sharing & 0.145 & 1 \\ \hline
\(w_{m}\) & Weight of shared information for midwives & 0.118 & 1 \\ \hline
\(x_{h}\) & Health payoff for healthy delivery & 0.457 & 100 \\ \hline
\(s_{i}[a_{i}]:s_{i}[a_{\neg i}]\) & Psuedo-count favouring honesty & 0.140 & 100:1 \\ \hline
\end{tabular}
\caption[Table caption text]{IQR of median moderate drinker signalling parameter sensitivity. \label{tab:sa_results_sig_iqr}}
\end{table}

\begin{table}[h!]
\center
\begin{tabular} {|l | l | l | l | l | l |}
\hline
Rule & \(\sigma^2\) & Nugget \(\sigma^2\) & Mean output & Total output variance & Code uncertainty & RMSSE \\ \hline
Lexicographic & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
Bayesian Payoff & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
Bayesian & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
\ac{CPT} & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
\end{tabular}
\caption[Table caption text]{IQR of median between groups IQR emulator statistics. \label{tab:sa_emulator_sig_iqr}}
\end{table}

\subsubsection{Between Groups IQR IQR}

\begin{table}[h!]
\center
\begin{tabular} {|l | l | l | l | l | l |}
\hline
Parameter & Description & Lexicographic & Bayesian Payoff & Bayesian & \ac(CPT) \\ \hline
\(p_{w}(m)\) & Proportion of moderate drinkers & 0.367 & 1 \\ \hline
\(p_{w}(l)\) & Proportion of light drinkers & 10.080 & 1 \\ \hline
\(p_{m}(m)\) & Proportion of moderate midwives & 6.715 & 1 \\ \hline
\(p_{m}(l)\) & Proportion of non-judgemental midwives & 43.942 & 1 \\ \hline
\(q_{w}\) & Probability of women sharing & 0.198 & 1 \\ \hline
\(w_{w}\) & Weight of shared information for women & 0.355 & 1 \\ \hline
\(q_{m}\) & Probability of midwives sharing & 0.145 & 1 \\ \hline
\(w_{m}\) & Weight of shared information for midwives & 0.118 & 1 \\ \hline
\(x_{h}\) & Health payoff for healthy delivery & 0.457 & 100 \\ \hline
\(s_{i}[a_{i}]:s_{i}[a_{\neg i}]\) & Psuedo-count favouring honesty & 0.140 & 100:1 \\ \hline
\end{tabular}
\caption[Table caption text]{IQR of median between groups IQR parameter sensitivity. \label{tab:sa_results_iqr_iqr}}
\end{table}

\begin{table}[h!]
\center
\begin{tabular} {|l | l | l | l | l | l |}
\hline
Rule & \(\sigma^2\) & Nugget \(\sigma^2\) & Mean output & Total output variance & Code uncertainty & RMSSE \\ \hline
Lexicographic & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
Bayesian Payoff & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
Bayesian & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
\ac{CPT} & 0.834 & 0.131 &  0.817 & 0.012 & 0.252 & 1.746 \\ \hline
\end{tabular}
\caption[Table caption text]{IQR of median between groups IQR emulator statistics. \label{tab:sa_emulator_iqr_iqr}}
\end{table}