### 2014-12-18 Conclusion.tex
More comments

%!TEX root = disclosure_game.tex
\section{Discussion}
\label{sec:discussion}
%reiterate contribution
%	bring up the simplicity -> chaos thing, talk about homophily
%	talk up the good (trends! yay!) and intro the bad -> no data
%
%	emphasise the 'nice story' -> future work
% ~ 1,500 words.

From a pragmatic perspective, the differing response characteristics of the decision rules are substantial and significant. The lack of a mechanism of homophily in the very simple rules leads to a high level of uncertainty in the overall dynamics, since even irrelevant information is taken at face value and to be equally valuable. By contrast, the belief structure of the \ac{CPT}, and Bayesian models represents a useful abstraction which allows the use of the relevant aspects of the information.  Naturally, incorporating homophily, by, for example weighting shared information by the type of the sharer, would represent a relatively trivial modification to the heuristic models. While to some extent this highlights the flexibility of the decision rule approach, it would of course sacrifice the parsimony of the model to a degree. This is an important consideration, given that part of the argument in favour of a decision theoretic approach to agent building lies in the minimal nature of the behavioural rules, which can be seen as occupying a middle ground between parameter and model.

One of the notable features of the results is the similarity in behaviour of the two classes of decision rule. To some extent this reflects poorly on the most complex rule, \ac{CPT}, which diverges only minimally in behaviour from the Bayesian model. To some extent this might be anticipated, in that the payoffs are clearly unrealistic, which limits the utility of the \ac{CPT} approach.  Additionally, work by \citet{Glockner2012} has shown that there is considerable variation in individual parameters for the decision model, whereas we have let them remain homogenous here. In the same vein, utility functions should arguably vary between individual agents, which could potentially be addressed by replacing the fixed payoffs used here with a distribution.  With this said, the significant increase in complexity, which entails both additional parameters and increases to simulation time may auger for a middle ground, particularly where elicitation of payoffs is impractical.  This, together with the variablility associated with the heuristic type decision rules speaks to a tradeoff between capturing reality, and replicating it.

Continuing the discussion of the issues raised by the representation of payoffs, the temporal aspect is significant, in that there is clearly a timing difference in payoffs, since while the potential social pain of disclosure is immediate, the health outcome comes only later. In light of this, that there is a known impact of time on perceived utility \citet{Thaler1981} suggests that incorporating some form of temporal discounting (e.g. exponential \citep{Samuelson1937}, or hyperbolic \citep{Ainslie1991}), or a decision model which explicitly treats intertemporal choice, such as the \ac{CPT} like model of \citet{Loewenstein1992}, is warranted.

As noted in section \ref{sub:sensitivity}, the impact of information sharing between midwives is surprisingly minimal, where it might be expected to play a more significant role in reality. A possible explanation for this lies in the implementation, which may place an excessive constraint on how much information midwives can share. The restriction to sharing only after a referral, together with the disparity in population sizes, and random allocation of appointments leads to midwives rarely having more than a single interaction with woman to pass on to their colleagues. Furthermore, because midwives are only informed of the true type if a referral occurs, they have an inherent myopia, since until they have evidence of deception they will not refer, with said evidence difficult to obtain without a referral.

In reality it might be anticipated that midwives would not withhold judgement so completely, and would pass on concerns about particular women to their colleagues, or that particularly dramatic stories would persist and be passed. This might be addressed by incorporating noisy type information \citep{Feltovich2002}, capturing the unintentional information transmitted during appointments, together with a relaxation of the assumptions about when information may be shared and more sophisticated model of information transmission in general. This also highlights an advantage of the \ac{BACCO} approach, in that it aids in diagnosing issues with model design by giving insight into parameters which are contributing innapropriately to variance in output.

\section{Conclusion}
\label{sec:conclusion}

The conclusions that can be drawn about the behaviours of real life women, and their midwives, are necessarily limited by the paucity of data available to validate the model. While qualitative trends offer some indication, they are clearly very limited in scope, and do not permit strong claims about the drivers of disclosure.  As such, further work will focus on applying the model to domains where validation data is more available, which will support a more comprehensive evaluation of the model discrepancy. 
With this said, the trends reported by \citet{Alvik2006}, and \citet{Phillips2007} are borne out by the model, and predictions from the two more complex rules suggest that encouraging information sharing between women may encourage disclosure, but at the expense of reducing accuracy. By contrast, if one takes the view that a Lexicographic model is a better approximation of real behaviour, then outcomes can best be influenced by controlling how far midwives punish their women socially. We would however suggest that there are better reasons than the outputs of a simulation for doing so.

More broadly, the results demonstrate the logistical feasibility, and its utility as a `tool for thinking', of an agent model grounded in decision theory. The results also make clear that deciding the operationalisation of the decision making is of key significance.

 % Disclosure is a widely applicable issue in health and social care and beyond, with examples ranging from a lack of help seeking behaviour noted in older men \citep{Smith2007a}, to the inverse scenario of innapropriate disclosure in social media \citep{christofides2009information}.