### 2014-12-31 Results.tex
More Springer jazz

%!TEX root = disclosure_game.tex
%Include the descent into chaos bit
\section{Results}
\label{sec:results}

\subsection{Qualitative Trends}
\label{sub:qt_results}

As shown in figure \ref{fig:honest_signals}, all four decision rules were able to reproduce both qualitative trends towards more disclosure as women experience more appointments \citep{Phillips2007}, and a greater tendency towards underreporting of consumption by heavier drinkers \citep{Alvik2006}.  Trends for all four rules are broadly similar, exhibiting a gradual increase across appointments which subsequently levels off. This levelling can in part be explained by the referral results (figure \ref{fig:referrals}), which show that the majority of drinkers are referred, even with substantial concealment. Referrals continue to occur, in the absence of honest signals, because drinkers are able to achieve a referral by masquerading as higher or lower types, dependent on how their initial beliefs are biased. Despite this the results suggest that a minority of risky drinkers will evade detection altogether, with no notable distinction between heavy and moderate types. Under these parameters, light drinkers always signal honestly and are never referred since there is no perceived advantage in doing so, and the evidence of deceptive signalling is insufficient to outweigh the biased priors of the midwives.

\begin{SCfigure}[H]
\begin{adjustbox}{center}\subfloat[Average fraction of population ever signalled honestly\label{fig:honest_signals}]{\includegraphics[width=78mm]{figures/honesty_plot}\end{adjustbox}{center}
}
\vskip
\begin{adjustbox}{center}\subfloat[Average fraction of population referred\label{fig:referrals}]{\includegraphics[width=78mm]{figures/ref_plot}}\end{adjustbox}
\caption{Qualitative trends after 1000 rounds, mean with 95\% confidence limit over 1000 runs}
\label{fig:qt_trends}
\end{SCfigure}

\subsection{Information Sharing}
\label{sub:sharing_results}

Introducing information sharing amongst women leads the behaviour of the decision rules to diverge markedly. Figure \ref{fig:honest_sharing} shows the proportion of women who have signalled honestly at least once by their final appointment, under four sharing conditions. Aside from the lexicographic decision rule, the general tendency is towards less honest signalling by heavy drinkers, but figure \ref{fig:ref_sharing} shows a slight increase in referrals for the Bayesian, and \ac{CPT} rules. For these decision models, this is because information sharing exacerbates the existing tendency of heavy drinkers to impersonate moderate drinkers, who behave more honestly as heavy drinkers become less so.

Particularly notable, is the decline in honest signalling by light drinkers visible in both heuristic type rules at the 0.25 level of \(q_{w}\) \& \(w_{w}\), and the associated increase in false positives. This arises because of the lack of homophilly in information sharing, as light drinkers become informed about negative outcomes associated with concealment, despite having nothing to conceal. The relatively high referral rates of drinkers heighten the effect further, because shared information becomes dominated by their experiences.

The relationship is not, however, entirely straighforward, in that increasing information sharing leads to greater variance between runs. A linear model was used to predict the between runs interquartile range of the average signal sent by moderate drinkers, with decision rule, and level of information sharing, together with the interaction between the two as predictors. The regression results were significant (\(F_{7,12}=25,p<2.9\times10^{-6}\)) with \(R^2=94\%\), and intercept 0.07. The only significant coefficients were for the interaction terms, which were 0.44 (\(p<0.05\)) for the Bayesian payoff rule, and 0.69 (\(p<0.005\)) for the lexicographic. This suggests that information sharing, for the heuristic style decision rules introduces considerable uncertainty to the model, which is explored further in the sensitivity analysis below.

\begin{figure}[H]
\begin{adjustbox}{center}
\includegraphics[width=119mm]{figures/honesty_sharing}
\end{adjustbox}
\caption{Impact of information sharing on trends in honest signalling after 1000 rounds, mean with 95\% confidence interval over 100 runs}
\label{fig:honest_sharing}
\end{figure}

\begin{figure}[H]
\begin{adjustbox}{center}
\includegraphics[width=119mm]{figures/ref_sharing}
\end{adjustbox}
\caption{Impact of information sharing on referral trends after 1000 rounds, mean with 95\% confidence interval over 100 runs}
\label{fig:ref_sharing}
\end{figure}

\subsection{Sensitivity Analysis}
\label{sub:sa_results}

In this section we present a brief overview of the sensitivity analysis, followed by selected results highlighting the global effect of changes to perceived payoffs and degree of bias towards honesty, as well as information sharing within women. The full results for the sensitivity analysis covering all sixteen emulators are available in appendix \ref{app:sensitivity_results}.

For the median signal choice of moderate drinkers, the results of the sensitivity analysis suggest that the proportion of light drinkers has a significant effect for all decision rules, accounting for 10\%, 38\%, 24\%, and 5\% of the variance in output for the Lexicographic, Bayesian Payoff, Bayesian, and \ac{CPT} rules respectively. For the Lexicographic rule, the overwhelming majority of variance in signalling behaviour is reflective of the prevalence of stigmatisation by midwives (44\% \(p_{m}(m)\), 7\% \(p_{l}(m)\), and a further 15\% for their interaction).  The proportions of midwives are also key drivers in group separation, and the between run IQR of both measures for this rule.

Perhaps surprisingly, variance attributable to information sharing between midwives is relatively low, with neither the weight nor probability accounting for more than 5\% of variance in any measure. While there are small contributions to variance in interaction with other parameters (e.g. 4\% to between groups IQR, for the interaction with the proportion of light drinkers under the Bayesian rule), this may suggest that the implementation is lacking, which we touch on in section \ref{sec:conclusion}.

\begin{figure}[H]
\begin{adjustbox}{center}\includegraphics[width=119mm]{figures/sharing_emulated_simulated}\end{adjustbox}
\caption{Median moderate drinker signal vs median between drinking type IQR for all decision rules}
\label{fig:outcome_plots}
\end{figure}

Figure \ref{fig:outcome_plots} gives a qualitative picture of both emulator quality, and the divergent response surfaces of the decision rules in response to variations in information sharing parameters. Emulator fit is clearly imperfect, but overall behaviour is qualitatively similar, with both emulated and simulated plots demonstrating separation in outcome space for the decision rules.

\begin{figure}[H]
\begin{adjustbox}{center}\includegraphics[width=119mm]{figures/unfixed_emu_sig_iqr}\end{adjustbox}
\caption{Emulated moderate drinker signal IQR in response to varying \(q_{w}\), and \(w_{w}\)}
\label{fig:emulated_sharing_iqr}
\end{figure}

Following from the suggestive results for information sharing introducing uncertainty (section \ref{sub:info_sharing}), figure \ref{fig:emulated_sharing_iqr} shows emulated points covering the parameter space in high resolution. These plots reflect the increase in uncertainty of outcome shown for the heuristic type rules, which is especially severe for the Bayesian payoff rule. They also suggest that the Bayesian decision rule is less stable under conditions where the weight of shared information is substantially higher than the probability of sharing. Also of interest is the large region of instability that can be discerned for the Bayesian rule under modest probability of information sharing and higher weights. This suggests that, perhaps unsurprisingly, that what information is shared is significant to outcomes.

For the \ac{CPT}, and Bayesian decision models, the interaction of bias towards honesty, and distinction between payoffs has a significant, and non-linear effect on instability, and separability of groups. Figure \ref{fig:emulated_payoff_group_iqr} shows the effects, and also highlights the tendency towards poor separability of groups for both the heuristic type decision rules. The response surface of the Bayesian payoff rule is slightly more nuanced than the simple Lexicographic rule. Figure \ref{fig:emulated_payoff_group_iqr} shows better separation, close to partial pooling\footnote{Pooling occurs when signallers of different types `pool' their signals, and one adopts the signals of another.} at high payoff distinction, with relatively modest honesty bias, which is reflected by the variance contributions of 11\% and 8\% respectively.  For the more complex rules, the general tendency is towards less pooling for higher values of both, but with pockets where full pooling\footnote{Indicating that all signallers types are using a single signal.} occurs.  The plots also suggest that the sensitivity of the \ac{CPT} rule is marginally greater, which is supported by the significant contribution to variance of close to 15\% for all measures of \(x_{h}\).

\begin{figure}[H]
\begin{adjustbox}{center}\includegraphics[width=119mm]{figures/unfixed_emu_payoff_honesty_group_iqr}\end{adjustbox}
\caption{Emulated between groups IQR in response to varying \(s_{i}[a_{i}]:s_{i}[a_{\neg i}]\), and \(x_{h}\)}
\label{fig:emulated_payoff_group_iqr}
\end{figure}