### 2015-06-15 simulation_schedule.tex

%!TEX root = disclosure_game.tex

\section{Simulation Schedule}
\label{app:sim_schedule}

In this section we provide psuedocode for a single simulation, which encompasses several games.

\begin{algorithm}
\begin{algorithmic}
\State $rounds \gets 1000, appointments \gets 12$
\State women \gets enqueue([], $w_{1},\ldots,w_{1000}$), midwives \gets [$m_{1},\ldots,m_{100}$]
\While{$rounds > 0$}
	\State players $\gets$ dequeue first 100 women
	\State referred $\gets$ []
	\State $i $\gets$ 1$
	\While{$i < 100$}
		\State woman $\gets$ players[i], midwife $\gets$ midwives[i]
		\State signal $\gets$ signal(woman)
		\State response $\gets$ respond(midwife)
		\State midwife's memory $\gets$ woman, signal, response
		\State woman's memory $\gets$ type(midwife), signal, response, $X_{w, s, r, type(woman), type(midwife)}$
		\State update woman's beliefs
		\If{$response = refer$}
			\State midwife's memory $\gets$ type(woman), $X_{m, s, r, type(woman)}$
			\State retrospectively update midwife's beliefs
			\State women $\gets$ enqueue(women, new woman)
			\State Add woman to referred
		\Else
			women $\gets$ enqueue(women, woman)
		\EndIf
	\EndWhile
	\For{woman in referred}
		\If{$random \lt p$}
			\For{player in women}

			\EndFor
		\EndIf
	\EndFor
\State $rounds \gets rounds - 1$
\EndWhile
\end{algorithmic}
\caption{Simulation schedule\label{alg:sim}}
\end{algorithm}