### 2015-06-19 disclosure_game.tex

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}
\usepackage[hyperfootnotes=false]{hyperref}
\hypersetup{bookmarksdepth=3}
% choose options for [] as required from the list
% in the Reference Guide

\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                            % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom
%\documentclass{article}
% \usepackage{tikz}
 \usepackage{adjustbox}
 \usepackage{multirow}
 \usepackage{algorithm}
 \usepackage{algpseudocode}
 \usepackage{ifthen}
 \usepackage{csquotes}
 \usepackage{amsmath}

\usepackage{comment}
 \usepackage{natbib}
 \setcitestyle{numbers,square}
% \usepackage{hyperref}
 \usepackage{graphicx}
 \usepackage[subrefformat=parens,labelformat=simple,caption=false,labelfont={bf}]{subfig}
 \usepackage[nolist,smaller]{acronym}
 \usepackage{sidecap}
% \usepackage{nomencl}
\makeindex

\begin{document}
\begin{acronym}[ALSPAC]
  \acro{FAS}{Foetal Alcohol Syndrome}
  \acro{CPT}{Cumulative Prospect Theory}
  \acro{PT}{Prospect Theory}
  \acro{NICE}{National Institute for Health and Care Excellence}
  \acro{IAPT}{Improving Access to Psychological Therapies}
  \acro{CMACE}{Centre for Maternal and Child Enquiries}
  \acroindefinite{CMACE}{the}{the}
  \acro{AUDIT}{Alcohol Use Disorders Identification Test}
  \acro{T-ACE}{Tolerance, Annoyance, Cut down, Eye-opener}
  \acro{RCOG}{Royal College of Obstetricians and Gynecologists}
  \acroindefinite{RCOG}{the}{the}
  \acro{AML}{acute myeloid leukemia}
  \acro{ADHD}{Attention Defecit Hyperactivity Disorder}
  \acro{PND}{Postnatal Depression}
  \acro{RCT}{randomised control trial}
  \acro{WHO}{the World Health Organisation}
  \acro{ESS}{Evolutionarily Stable Strategy}
  \acrodefplural{ESS}[ESS]{Evolutionarily Stable Strategies}
  \acro{MCMC}{Markov chain monte carlo methods}
  \acro{DU}{Discounted Utility}
  \acro{AIC}{Akaike information criterion}
  \acro{ANOVA}{analysis of variance}
  \acro{ALSPAC}{the Avon Longitudinal Study of Parents and Children}
  \acro{GEM}{Gaussian Emulation Machine}
  \acrodefplural{GEM}{Gaussian Emulation Machines}
  \acro{GEM-SA}{Gaussian Emulation Machine for Sensitivity Analysis}
  \acro{JIT}{Just-in-time}
  \acro{ABM}{Agent Based Modelling}
  \acro{ABMs}{Agent Based Models}
  \acro{CPT}{Cumulative Prospect Theory}
  \acro{FFH}{Fast and Frugal Heuristic}
  \acro{OFC}{orbito-frontal cortex}
  \acro{BACCO}{Bayesian Analysis of Computer Code Outputs}
  \acro{IQR}{interquartile range}
\end{acronym}

\title*{Deciding to Disclose}
\subtitle{A Decision Theoretic Agent Model of Pregnancy and Alcohol Misuse}

\author{Jonathan Gray, Jakub Bijak, and Seth Bullock}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Jonathan Gray \at University of Southampton, Southampton, United Kingdom \email{j.gray@soton.ac.uk}
\and Jakub Bijak \at University of Southampton, Southampton, United Kingdom \email{j.bijak@soton.ac.uk}
\and Seth Bullock \at University of Southampton, Southampton, United Kingdom \email{sgb@ecs.soton.ac.uk}}

\maketitle

%*************************
% Acronyms
%*************************
%\phantomsection
%\refstepcounter{dummy}
%\pdfbookmark[1]{Acronyms}{acronyms}
%\markboth{\spacedlowsmallcaps{Acronyms}}{\spacedlowsmallcaps{Acronyms}}
%\section*{Acronyms}

%\refstepcounter{dummy}
%\pdfbookmark[1]{Glossary}{glossary}
%\printglossary[type=main]

%\printnomenclature{}

%Assuming this is for JASSS.. Need to make background shorter.
%Aim for ~ 7-8,000 words overall.

\input{abstract_short}
\acresetall

\input{Introduction} %(half page)

%\input{Background}

\input{Model}
\input{agent_models}

\input{Methods}

\input{Results}

\input{Conclusion}
\input{Acknowledgements}

\bibliography{library,library_addendum}
\bibliographystyle{spbasic}
\input{SensitivityAnalysis}
\input{model_description}
\input{simulation_schedule}
\input{agent_examples}
\end{document}