### 2017-01-16 simulation_schedule.tex

%!TEX root = disclosure_game.tex

\chapter{Disclosure Game Simulation Schedule}
\label{app:sim_schedule}

In this section we give the step by step process for a single run of the disclosure game simulation.

\begin{enumerate}
\item Generate 1000 women, and place them in a queue
\item Generate 100 midwives.
\item For each round of the game
\begin{enumerate}
	\item Take 100 women from the queue
	\item Pair each one with a random midwife
	\item For each pair
	\begin{enumerate}
		\item The woman sends a signal
		\item The midwife refers or not based on the signal
		\item The woman is informed of her payoff, the midwife's type, and whether she is referred
		\item The woman updates her beliefs
		\item The midwife stores the game in their memory
		\item If the woman is referred
		\begin{enumerate}
			\item The midwife is informed of the woman's true type
			\item The midwife retrospectively updates their beliefs using the true type, and memories of any games with this woman
			\item The midwife is now eligible to share their memories of games with this woman
		\end{enumerate}
	\end{enumerate}
	\item Women who have not been referred or had their baby, join the back of the queue
	\item New women are generated to replace those referred, or delivered
	\item The new women are added to the back of the queue
	\item For each referred or birthed woman
	\begin{enumerate}
		\item With probability p, her memory of games is shared with the active women
		\item She is removed from simulation
	\end{enumerate}
	\item The active women update their beliefs
	\item For each midwife with information to share
	\begin{enumerate}
		\item With probability p, their memory of games with the referred woman is shared
		\item The memory is no longer eligible to be shared
	\end{enumerate}
	\item The midwives update their beliefs
\end{enumerate}
\end{enumerate}